Self-reported LLM usage and outcomes on a data science project: Evidence from a Canadian undergraduate data science course

### 2025-03-14 paper/paper.qmd
Reconcile Lucas updates with mine

---
title: "Self-reported LLM usage and results on a data science project: Evidence from a Canadian undergraduate data science course"
author: 
  - Rohan Alexander
  - Luca Carnegie
  - Nathalie Moon
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement. We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions. This research is currently under review by the University of Toronto's Research Ethics Board. Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing. Comments can be sent to: rohan.alexander@utoronto.ca."
date: today
date-format: long
abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting. We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark. Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists. Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
format: pdf
number-sections: true
table-of-contents: false
bibliography: references.bib
---

@@

Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner. These principles are upheld in various ways, through ongoing education, adherence to professional norms and, and implementation of reproducible workflows.  To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.  Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit. These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.

Like all new tools LLMs have created both excitement and apprehension. ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation. LLMs have quickly been put to use in both industry and academia. By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.

@@

The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building]. One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python. Another is writing, which is primarily done when communicating results. The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from their already-demonstrated capability in those competencies in adjacent professional fields.

@@

At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well. Teachers were also concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance. Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.

@@

To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage. Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general. This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto. By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.

The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students. @sec-model specifies a model used to investigate the relationship. @sec-results describes and analyzes the model's results. @sec-discussion discusses the implications of the findings for data science education and future research and practice at the intersection of LLMs and trustworthy data science.

@@

Data was collected from students through an optional end-of-course survey. [Appendix -@sec-survey-details] details the questions asked in the survey. Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation. Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper. The responses were anonymized by removing any personal references to the students themselves including, names, emails, student numbers, and GitHub links.

Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.

@@

There were 146 responses to the survey. Of these, 119 respondents provided authorization for their data to be collected and used. Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained. Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses. Finally 7 of those respondents did not provide a usable GPA response. This leaves 93 respondents that were of use and were merged based on student name.

All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1). That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out. That respondent is included in the analysis dataset, but without them, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.

@@

There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2). The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55. One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments. Self-reported GPAs also introduce the possibility of reporting bias. For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.

Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist). The course's prerequisite two-course sequence is typically completed by students in their second year, which makes it difficult for students to take this course earlier than their 3rd year.

@@

# so that "Very different from me" is Red - better visuals
set1_palette <- c("#4DAF4A", "#377EB8", "#E41A1C")

writing_perceptions_long |>
  ggplot(aes(x = response, fill = response)) +
  geom_bar(position = "stack", stat = "count") +
  facet_wrap(vars(statements)) +
  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
  scale_fill_manual(values = set1_palette) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.position = "bottom")

@@

coding_perceptions_long |>
  ggplot(aes(x = response, fill = response)) +
  geom_bar(position = "stack", stat = "count") +
  facet_wrap(vars(statements)) +
  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
  scale_fill_manual(values = set1_palette) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

@@

The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1). Though respondents' self-perceptions around writing and coding are varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2). Most respondents who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.

@@

To understand the role of LLMs in learning, students identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1). Technical questions and explaining concepts were the two top use cases among students in the course. More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions. Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing; this could be similar to the attitudes expressed by students surveyed in @Cahill_2024.

@@

> Many students view AI as a helpful supplementary tool, comparing it to resources like Google or calculators. They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks. However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking. Students emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively. Some argue that learning to use AI effectively is a valuable skill for future careers. Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information. Overall, most students support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.

@@

We find a varied extent of self-reported LLM usage (@tbl-llm-usage). 41 respondents were classified as having made extensive use of LLMs, while 28 were classified as having made somewhat use. 31 respondents were classified as having made minimal or no use of LLMs. In the analysis dataset we combine those two classifications because only 8 respondents were classified as having minimal usage.

@@

The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1). The overall mean was 78% and standard deviation was 17 percentage points. However there was considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).

@@

The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage. Students received their result on the final paper as a proportion (a 0-100% mark), so we use a Bayesian zero-one-inflated beta regression to best fit the distribution in this interval. We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed  (heteroskedastic) datasets, which we can see in the asymmetrical histogram of the distribution of marks.

Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.

We briefly describe the model that we use, which follows @kurztotherescue, which was adapted from @ospina. No students got zeroes, though some got full marks (100%) on the paper, so we adapt the regression to one-inflated beta model as described by @heisstotherescue to account for this.

Model diagnostics are included in [Appendix -@sec-model-details].

The model uses a one-inflated beta regression approach to analyze the relationship between student LLM usage, GPA, and final paper marks. Let $Y_i$ represent the mark for student $i$, modeled as:

$$Y_i \sim \text{Zero-One-Inflated Beta}(\mu_i, \phi_i, \pi_{zoi_i}, \pi_{coi_i})$$

where:

@@

The model components are specified as:

@@

The "llm_usage" predictor is categorical, with students' self-reported LLM usage classified as "Extensive" (reference level), "Somewhat", or "None or minimal". GPA is a continuous variable representing students' self-reported grade point averages. The model is implemented using the brms package in R, which provides a Bayesian framework for fitting these regression models.

The expected relationship between LLM usage and final mark is unclear, given the modelling strategy and variables collected. It could be that stronger students used LLMs in a more sophisticated way or that they did not need to use them. However, we can expect the relationship between GPA and final mark to be positive.

@@

Our key results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.

@tbl-model-results and @fig-model-results present the key results from two models estimating the association between LLM usage, GPA and final paper mark. The first model includes only LLM usage, while the second model also includes self-reported GPA as a predictor.

```{r tbl-model-results}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Selected Coefficient estimates and mean absolute deviation (MAD)

fit1 <-
  readRDS(file = here("models/fit_one_inflated.rds"))

fit2 <-
  readRDS(file = here("models/fit_one_inflated_gpa.rds"))

modelsummary(
  list(
    "Base model" = fit1,
    "Self-reported GPA Included" = fit2
  ),
  statistic = "mad",
  fmt = 2,
  output = "kableExtra",
  kable_options = list(
    booktabs = TRUE,
    linesep = ""
  ),
  coef_map = c(
    "b_llm_usageSomewhat" = "LLM usage: Somewhat",
    "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
    "b_what_is_your_gpa" = "GPA", 
    "b_zoi_Intercept" = "Zero-one inflation intercept"
  )
)
```

```{r fig-model-results}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
modelplot(
  list(
    "Base model" = fit1,
    "Including self-reported GPA" = fit2
  ),
  conf_level = 0.9,
  coef_map = c(
    "b_llm_usageSomewhat" = "LLM usage: Somewhat",
    "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
    "b_what_is_your_gpa" = "GPA"
  ),
  coef_omit = NULL
) +
  labs(x = "") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```

@@

The overall fit of the model improves when GPA is included. There is an solid increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics. This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).

@@

1. LLMs could help enhance the trustworthiness of data science, but not without considerable further development, testing and integration in the data science process.

2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help, and

3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science in particular.

@@

There are several, substantial weaknesses of this study. The foundational one is that we used observational data, much of which was self-reported. There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class. A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.

@@

modelplot(
  list(
    "Base model" = fit1,
    "Including self-reported GPA" = fit2
  ),
  conf_level = 0.9,
  drop = c("phi", "b_zoi_Intercept")) +
  labs(x= "") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

Next, we use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.

@@

>>>>

---
title: "Self-reported LLM usage and outcomes on a data science project: Evidence from a Canadian undergraduate data science course"
author: 
  - Rohan Alexander
  - Luca Carnegie
  - Nathalie Moon
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement. We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions. This research is was approved by the University of Toronto's Research Ethics Board as Protocol 47725. Comments can be sent to: rohan.alexander@utoronto.ca."
date: today
date-format: long
abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which self-reported LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting. We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark. Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists. Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
format: pdf
number-sections: true
table-of-contents: false
bibliography: references.bib
---

<!-- Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing.  -->

@@

Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner. These principles are upheld in various ways, through ongoing education, adherence to professional norms and implementation of reproducible workflows. To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions. Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit. These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.

Like all new tools LLMs have created both excitement and apprehension. ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation. LLMs have quickly been put to use in both industry and academia. By now many people, especially educators and students, have some experience with LLMs in both personal and professional contexts.

@@

The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building]. One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python. Another is writing, which is primarily done when communicating results. The potential for LLMs to positively affect students' academic performance in data science follows from their already-demonstrated capability in those competencies in adjacent professional fields.

@@

At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that LLMs could help creativity, provide academic support when teachers were unavailable, and model certain types of writing well. But teachers were concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance. Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.

@@

To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage. Specifically, we examine the relationship between the mark a student got on a final paper and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general. This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto. By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.

The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data, self-reported LLM usage, and final marks. @sec-model specifies a model used to investigate the relationship. @sec-results describes and analyzes the model's results. @sec-discussion discusses the implications of the findings for data science education and future research and practice at the intersection of LLMs and trustworthy data science.

@@

Data were collected from students through an optional end-of-course survey. [Appendix -@sec-survey-details] details the questions asked in the survey. Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation. Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper. The responses were then anonymized by removing any personal references to the students themselves including, names, emails, student numbers, GitHub links, and summarizing free-text responses.

Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the 
`arrow` [@arrow], 
`here` [@here], 
`janitor` [@janitor], 
`readxl` [@readxl], and 
`tidyverse` [@tidyverse] 
packages.

@@

There were 146 responses to the survey. Of these, 119 respondents provided authorization for their data to be collected and used. Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained. Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses. Finally 7 of those respondents did not provide a usable GPA response.

All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1). That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out. After that respondent is removed from the analysis the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes. This leaves 92 respondents that were of use and were merged based on student name.

@@

There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2). The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55. One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments. While there was no reason for the respondents to not report the truth, self-reported GPAs also introduce the possibility of reporting bias. For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.

Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist). The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.

@@

writing_perceptions_long |>
  ggplot(aes(x = response, fill = response)) +
  geom_bar(position = "stack", stat = "count") +
  facet_wrap(vars(statements)) +
  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.position = "bottom")

@@

coding_perceptions_long |>
  ggplot(aes(x = response, fill = response)) +
  geom_bar(position = "stack", stat = "count") +
  facet_wrap(vars(statements)) +
  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

@@

The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1). Though respondents' self-perceptions around writing and coding were varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2). Most respondents who selected "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.

@@

To understand the role of LLMs in learning, respondents identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1). Technical questions and explaining concepts were the two top use cases among respondents. More than half of respondents also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions. Just less than half reported that they used it to write paper content, which could suggest that respondents do not feel confident using it to improve writing.

@@

> Many respondents view AI as a helpful supplementary tool, comparing it to resources like Google or calculators. They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks. However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking. Respondents emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively. Some argue that learning to use AI effectively is a valuable skill for future careers. Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information. Overall, most respondents support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.

As at 14 August 2024, the default Anthropic setting is that they do not use input data for training and so these student responses should have entered future training datasets [@anthropic].

@@

We find a varied extent of self-reported LLM usage (@tbl-llm-usage). 38 respondents were classified as having made extensive use of LLMs, while 26 were classified as having made somewhat use. 28 respondents were classified as having made minimal or no use of LLMs. In the analysis dataset we combine those two classifications because only a handful of respondents were classified as having minimal usage.

@@

The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1). The overall mean was 78% and standard deviation was 17 percentage points. However there were considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).

@@

The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage. Students received their result on the final paper as a percentage between 0% and 100%. There were no students who got 0%, but 9 students got 100%. As such, we use one-inflated beta regression [@Ospina2012]. Beta regression is commonly used when data are between 0 and 1 and is focused on estimating the mean and variance of the beta distribution. One approach is to consider a one-inflated distribution which is a mixture of focusing on those responses that were 1, and the usual Beta distribution aspect.

The main variable of interest, `llm_usage` is a categorical variable with three possible values: "Extensive", "Somewhat", "None or minimal". We use "None or minimal" as the reference level and so the results are in relation to that level of self-reported LLM usage. Self-reported GPA, `what_is_your_gpa`, is a number between 1.9 and 4 with one decimal place.

Define $y_i$ as the percentage that student $i$ received on the final paper:

$$y_i \sim \mbox{Zero-One-Inflated Beta}\left(\mu_i, \phi_i, \mbox{zoi}_i, \mbox{coi}\right)$$

Where:

@@

The model components are:

@@

We estimate the model and explore the results using the statistical programming language `R` [@citeR], and the 
`brms` [@brms], and 
`modelsummary` [@modelsummary] 
packages. Our code to adjust for the fact that there were no students who got 0% follows @buerkner and @Heiss2021. Model diagnostics are included in [Appendix -@sec-model-details]. Finally, the piecewise approach is a common one for when there may be something different about the 0s or 1s, compared with the other responses. Another approach for dealing with a situation in which there are exact 0s or 1s is to slightly perturb the data. In our case, there is not much difference between getting 99% on a paper and getting 100%. When we estimated that perturbed data model using Beta regression the results were similar.

@@

Our results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.

```{r tbl-model-results}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Selected coefficient estimates and mean absolute deviation (MAD)

modelsummary(
  list(
    "Base model" = fit1,
    "Self-reported GPA included" = fit2
  ),
  statistic = "mad",
  fmt = 2,
  output = "kableExtra",
   kable_options = list(
     booktabs = TRUE,
     linesep = ""
   ),
   coef_map = c(
     "b_llm_usageSomewhat" = "LLM usage: Somewhat",
     "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
     "b_what_is_your_gpa" = "GPA", 
     "b_zoi_Intercept" = "Zero-one inflation intercept"
   )
)
```

```{r fig-model-results}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals

modelplot(
  list(
    "Base model" = fit1,
    "Including self-reported GPA" = fit2
  ),
  conf_level = 0.9,
 coef_map = c(
     "b_llm_usageSomewhat" = "LLM usage: Somewhat",
     "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
     "b_what_is_your_gpa" = "GPA"
   ),
   coef_omit = NULL
 ) +
   labs(x = "") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@tbl-model-results and @fig-model-results present the results from two models estimating the association between LLM usage, GPA and final paper mark. The first model includes only LLM usage, while the second model also includes self-reported GPA.

@@

The overall fit of the model improves when GPA is included. There is an increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics. This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).

@@

1. LLMs could help enhance the trustworthiness of data science, but not without considerable further development, testing and integration in the data science process.
2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help, and 
3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science in particular.

@@

There are several substantial weaknesses of this study. The foundational one is that we used observational data, much of which was self-reported. There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class. A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.

@@

modelplot(
  list(
    "Base model" = fit1,
    "Including self-reported GPA" = fit2
  ),
  conf_level = 0.9,
  drop = c("phi", "b_zoi_Intercept")
) +
  labs(x = "") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

@@

We use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
